---
title: "Advanced R - Chapter 2 | Names and values"
output: 
  html_document: 
    toc: yes
    code_folding: show
editor_options: 
  chunk_output_type: console
---

\
\
Advanced R - By Hadley Wickham (https://adv-r.hadley.nz/index.html)
\
\

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      fig.width = 14,
                      fig.height = 10)

# same seed throughout the book
set.seed(1014)
```





# Chapter 2 - Names and values
__________________________________________________________________________________________________

**NOTE:** RStudio may give different results. Thus it is recommended to use the Rterminal instead of Rmarkdown... I use Rmarkdown only for documentation purposes.
__________________________________________________________________________________________________

```{r}
# load packages
library(lobstr)
```


## 2.2 Binding basics

```{r}
x <- c(1, 2, 3)
```

"Create an object named 'x', containing the values 1, 2, and 3" is unfortunately a simplification. In the background R is creating an object (in this case a vector with the values 1, 2, and 3) and R is binding that object to a name (in this case 'x'). '<-' creates a binding from the name on the left-hand side to the object on the right-hand side (in this sense, the sign '<-' is a bit counterintuitive '->' would be better). 
=> Names as references to values.
In the following code the value 'c(1, 2, 3)' is not being copied. Instead you get another binding to the already existing object.

```{r}
y <- x

# check: same memory address
obj_addr(x) == obj_addr(y)
```


### 2.2.1 Non-syntactic names

```{r}
?Reserved
```

If you want to create names that do not align with these rules you can use '``' (backticks).$


### 2.2.2 Exercise

**Q1**

Explain the relationship between a, b, c and d in the following code:

```{r}
a <- 1:10
b <- a
c <- b
d <- 1:10
```

=> a, b, and c refer to the same object (point to the same object). R creates multiple bindings to the same object. d has the same value 'c(1:10)' but another binding.

```{r}
# same memory addresss/binding
obj_addr(a) == obj_addr(b) 
obj_addr(a) == obj_addr(c)

# different memory address/binding
obj_addr(a) == obj_addr(d)
```


**Q2**

The following code accesses the mean function in multiple ways. Do they all point to the same underlying function object? Verify this with lobstr::obj_addr().

```{r }
mean
base::mean
get("mean")
evalq(mean)
match.fun("mean")
```

=> Yes, they do:

```{r}
obj_addr(mean)
obj_addr(get("mean"))
obj_addr(evalq(mean))
obj_addr(match.fun("mean"))
```


**Q3**

By default, base R data import functions, like read.csv(), will automatically convert non-syntactic names to syntactic ones. Why might this be problematic? What option allows you to suppress this behaviour?

=> By default R 'corrupts' data (names are often data). This is problematic because there is no notification of the transformation. We can use 'check.names = FALSE' to suppress this behaviour. 


**Q4**

What rules does make.names() use to convert non-syntactic names into syntactic ones?
=> see ?make.nam


**Q5**

I slightly simplified the rules that govern syntactic names. Why is .123e1 not a syntactic name? Read ?make.names for the full details.
=> because it's a double
```{r}
.123e1
typeof(.123e1)
```



## 2.3 Copy-on-modify

```{r}
# same memory adress
x <- c(1, 2, 3)
y <- x


# copy-on-modify
y[[3]]
y[[3]] <- 4

x
y


# R created a new object and rebound 'y' to that new object.
obj_addr(x)
obj_addr(y)
```

=> copy-on-modify OR R objects are unchangeable or immutable.


### 2.3.1 tracemem()

You can use tracemem() to track the copy-on-modify behaviour.

```{r}
x <- c(1, 2, 3)
tracemem(x)

y <- x
y[[3]] <- 4L

obj_addr(x)
obj_addr(y)
```

If you modify y again, it will not be copied, because the new object only has a single name bound to it.

```{r}
y[[3]] <- 5L

obj_addr(y)

# trun off tracking
untracemem(x)
```


### 2.3.2 Function calls

The same rules for copying also apply for function calls:

```{r}
f <- function(a) {
  a
}

x <- c(1, 2, 3)
tracemem(x)

# create new object with same values
z <- f(x)
# no copy here!

untracemem(x)
```

Once f() completes, x and z will point to the same object. The 'original' object ('x') never gets copied because it never gets modified. If f() did modify x, R would create a new copy, and then z would bind that object.

```{r}
f <- function(a) {
  a[[3]] <- 4
}

x <- c(1, 2, 3)
tracemem(x)

# copy-on-modify!
z <- f(x)

untracemem(x)
```


### 2.3.3 Lists

Not just names (i.e. variables) that point to values; elements of lists do too:

```{r}
l1 <- list(1, 2, 3)
```

But lists are a bit more complex. Instead of storing the values itself, it stores references to them. This is important, when we modify a list:

```{r}
l2 <- l1
l2[[3]] <- 4
```

Like vectors, lists use copy-on-modify behaviour; the original list is left unchanged, and R creates a modified copy. This, however, is a shallow copy: the list object and its bindings are copied, but the values pointed to by the bindings are not. The opposite of a shallow copy is a deep copy where the contents of every reference are copied.

```{r}
# shallow copy
ref(l1, l2)
```


### 2.3.4 Data frames

Data frames are lists of vectors. Copy-on-modify has important consequences:

```{r}
d1 <- data.frame(
  x = c(1, 5, 6),
  y = c(2, 4, 3)
)

# If you modify a **column**, only that column needs to be modified. The others will still point to their original references:
d2 <- d1
d2[,2] <- d2[,2] * 2

ref(d1, d2)

# If you modify a **row** every columns needs to be copied:
d3 <- d1
d3 <- d3[1, ] * 3

ref(d1, d3)
```


### 2.3.5 Character vectors

R uses a **global string pool** where each element of a character vector is a pointer to a unique string in the pool.

```{r}
x <- c("a", "a", "abc", "d")

ref(x, 
    character = TRUE)
```


### 2.3.6 Exercises

**Q1**

Why is tracemem(1:10) not useful?

```{r}
tracemem(1:10)
```

=> the object has no name. So therefore it is not very useful to track it (no copy-on-modfy possible...)


**Q2**

Explain why tracemem() shows two copies when you run this code. Hint: carefully look at the difference between this code and the code shown earlier in the section.

```{r}
x <- c(1L, 2L, 3L)
tracemem(x)

x[[3]] <- 4

untracemem(x)
```

=> because we assign an element of type double to a list of type integer. This triggers the copy-on-modify. 
But we can easily avoid this:

```{r}
# all integer
x <- c(1L, 2L, 3L)
tracemem(x)

x[[3]] <- 4L


# all double
x <- c(1, 2, 3)
tracemem(x)

x[[3]] <- 4L
```


**Q3**

Sketch out the relationship between the following objects:

```{r}
# vector
a <- 1:10
# list
b <- list(a, a)
# list
c <- list(b, a, 1:10)
```

memory adresses/bindings

```{r}
# 3 different memory addresses/bindings
obj_addr(a)
obj_addr(b)
obj_addr(c)
```

references

```{r}
# a contains one references to the values 1:10
ref(a)
# b contains a list of two references to the same address as a
ref(b)
# c contains a list of b (containing two references to a), a (containing the same reference again) and a reference pointing to a different address containing the same value (1:10).
ref(c)
```


**Q4**

What happens when you run this code:

```{r}
x <- list(1:10)
x[[2]] <- x
```

The name x binds to a list object. This object contains a reference to the integer vector 1:10.

```{r}
x <- list(1:10)
obj_addr(x)
ref(x)
```

Copy-on-modify takes place! x is assigned to an element of itself => the list is copied to a new address in memory

```{r}
tracemem(x)

x[[2]] <- x
obj_addr(x)
ref(x)
```

x is a list of 2: the first element contains integers (1:10). The second is a list containing integers (1:10).

```{r}
str(x)
```


## 2.4 Object size

```{r}
obj_size(letters)
obj_size(LETTERS)

obj_size(ggplot2::diamonds)
```

Remember: Since the elements of a list are references to values, the size of a list might be much smaller than you expect:

```{r}
x <- runif(1e6)
obj_size(x)

y <- list(x, x, x)
obj_size(y)

obj_size(x, y)
```

R only stores the first and last value of a vector (this is called ALTREP = alternative representation)

```{r}
obj_size(1:3)
obj_size(1:1000)
obj_size(1:1e10)
```

Similarly to strings (because of the global string pool):

```{r}
banana <- "bananas bananas bananas"
obj_size(banana)

obj_size(rep(banana, 100))
```


### 2.4.1 Exercises

**Q1**

In the following example, why are object.size(y) and obj_size(y) so radically different? Consult the documentation of object.size().

```{r}
y <- rep(list(runif(1e4)), 100)

object.size(y)
# 8'005'648 bytes
obj_size(y)
# = 80'900 bytes
```

=> object.size() does not account for shared elements within lists... "Sizes of objects using a compact internal representation may be over-estimated." If you take a closer look, you notice, that a lot of elements are shared...

```{r}
str(y)
purrr::map(y, ~identical(., y[[1]]))
```



**Q2**

Take the following list. Why is its size somewhat misleading?

```{r}
funs <- list(mean, sd, var)
obj_size(funs)
```

=> all these functions are already included...


**Q3**

Predict the output of the following code:

```{r}
# base size
a <- runif(1e6)
obj_size(a)

# both lists contain references to the same memory address... No additional memory for the second list element.
b <- list(a, a)
obj_size(b)
obj_size(a , b)

ref(a, b)

# copy-on-modify => new address in memory for the first element. The second element still refers to the same address as a. Size of b doubles (2 different elements now). The size of a and b is equal to the size of b alone, because a is part of b...
b[[1]][[1]] <- 10
obj_size(b)
obj_size(a, b)

ref(a, b)

# copy-on-modify => new address for the second element too! So no element points to the same address as a anymore (2 new and separated addresses in memory...). Altogether we have 3 different vectors now => 3 * 8MB = 24MB
b[[2]][[1]] <- 10
obj_size(b)
obj_size(a, b)

ref(a, b)
```


## 2.5 Modify-in-place

=> Modifying an R object usually creates a copy. But there are 2 exceptions:

- Objects with a single binding get a special performance optimisation.
- Environments, a special type of object, are always modified in place.

### 2.5.1 Objects with single binding

```{r}
v <- c(1, 2, 3)
tracemem(v)

v[[3]] <- 4
```

for-loops are considered slow in R. That slowness is often caused by creating a copy with every iteration of the loop.

```{r}
x <- data.frame(
  matrix(
    runif(5 * 1e4), 
    ncol = 5
  )
)

str(x)

# calculate the medians for every element
medians <- vapply(x, median, numeric(1))

# substract the median from each column 
# (more precisely: form each element in each column)

start_time <- Sys.time()
for(i in seq_along(medians)) {
  
  x[[i]] <- x[[i]] - medians[[i]]
  
}
end_time <- Sys.time()
glue::glue("{end_time - start_time}")
```

What is going on in the background?

```{r}
tracemem(x)

for(i in seq_along(medians)) {
  
  x[[i]] <- x[[i]] - medians[[i]]
  
}

untracemem(x)
```

Each iteration copies the data frame twice (**NOTE:** Hadley speaks of three times. I was not able to reproduce that(not within Rmarkdown or the Rterminal...).)


### 2.5.2 Environments

Environments are always modified-in-place. This property is sometimes described as **reference semantics** because when you modify an environment all existing bindings to that environment continue to have the same reference.

```{r}
e1 <- rlang::env(a = 1, b = 2, c = 3)
e2 <- e1
```

If we change a binding, the environment is modified in place:

```{r}
e1$c <- 4
e1$c

e2$c
```

Environments can contain themselves (unique property of environments)

```{r}
e <- rlang::env()
e$self <- e

ref(e)
```


### 2.5.3 Exercises

**Q1**

Explain why the following code doesn’t create a circular list.

```{r}
x <- list()
x[[1]] <- x
```

Here copy-on-modify prevents the creation of a circular list.

```{r}
x <- list()
obj_addr(x)

tracemem(x)
# copy-on-modify
x[[1]] <- x

# new memory address of the copied object
obj_addr(x)
# old memory address
obj_addr(x[[1]])
```


**Q2**

Wrap the two methods for subtracting medians into two functions, then use the bench package to carefully compare their speeds. How does performance change as the number of columns increase?

```{r}
# create random data frame generator
create_random_df <- function(nrow, ncol) {
  
  # create random matrix
  random_matrix <- matrix(
    runif(nrow * ncol), 
    nrow = nrow
  )
  
  # transform to data frame
  as.data.frame(random_matrix)
}

create_random_df(2, 2)
```

```{r}
# create two different approaches to subtract medians

### 1) same for-loop as above...
subtract_df <- function(x, medians) {
  
  for (i in seq_along(medians)) {
    
    x[[i]] <- x[[i]] - medians[[i]]
  }
  
  return(x)
  
}

### 2) different approach
subtract_list <- function(x, medians) {
  
  x <- as.list(x)
  x <- subtract_df(x, medians)
  list2DF(x)
  
}
```

```{r}
# create median function
benchmark_medians <- function(ncol) {
  
  # create data frame and calculate medians
  df <- create_random_df(nrow = 1e4, ncol = ncol)
  medians <- vapply(df, median, numeric(1), USE.NAMES = FALSE)
  
  # evaluate via bench::mark()
  bench::mark(
    "data frame" = subtract_df(df, medians),
    "list" = subtract_list(df, medians),
    time_unit = "ms",
    iterations = 50
  )
  
}

benchmark_medians(1)
```

Test!

```{r}
# cretea container for results
results <- bench::press(
  ncol = c(1, 10, 50, 100, 250, 
           300, 400, 500, 750, 
           1000),
  benchmark_medians(ncol)
)

library(ggplot2)
theme_set(theme_minimal())


ggplot(results,
       aes(ncol, median, 
           color = as.character(expression))) +
  geom_point(size = 2) +
  geom_smooth() +
  labs(x = "Number of Columns",
       y = "Execution Time (ms)",
       colour = "Data Structure") +
  theme(legend.position = "top")
```


**Q3**

What happens if you attempt to use tracemem() on an environment?

=> not very meaningful => environments are always modified-in-place!


## 2.6 Unbinding and the garbage collector

Use the garbage collector (gc()) frees up memory by deleting R objects that are no longer used, and by requesting more memory from the operating system if needed. 
NOTE: actually there is no need to call gc() yourself. R will handle that for you...

```{r}
gc()

# wrapper around gc() => tells you how much memory is beeing used
lobstr::mem_used()
```

