---
title: "TidyTemplate"
date: 2021-06-25
output: html_output
editor_options: 
  chunk_output_type: console
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
```

# Load the weekly Data

Dowload the weekly data and make available in the `tt` object.

```{r Load}
library(malariaAtlas)

pr <- malariaAtlas::getPR(ISO = "KEN", species = "BOTH") # data only for Kenya
pr

class(pr)

# autoplot-function (malariaAtlas) only works with objects of class pr.points/data.frame
# PR = parasite rate (prevalence)
# for more information: https://github.com/malaria-atlas-project/malariaAtlas
autoplot(pr)
```



```{r}
kenya_pr <- as_tibble(pr)
kenya_pr %>% 
  View()

kenya_pr %>% 
  count(year_start) %>% 
  View()


kenya_pr %>% 
  group_by(year_start) %>% 
  summarise(examined = sum(examined),
            positive = sum(positive),
            studies = n()) %>% 
  mutate(pr = positive / examined) %>% 
  ggplot(aes(year_start, pr)) +
  geom_line()
# this line graph is confounded by the number of studies conducted each year (how much financing did they get?)

# sort of a map
kenya_pr %>% 
  ggplot(aes(longitude, latitude, color = year_start)) +
  geom_point() 
# why are there so many NAs in year_start? => "no permission to release data"
kenya_pr %>% 
  filter(is.na(year_start)) %>% 
  View()
```

```{r}
# filter out the NAs
kenya_pr <- kenya_pr %>% 
  filter(!is.na(pr))

kenya_pr %>% 
  ggplot(aes(longitude, latitude, color = year_start)) +
  borders(database = "world", regions = "Kenya") +
  geom_point() +
  coord_fixed()


kenya_pr %>% 
  arrange(pr) %>% # plot the highest points last! => very nice
  ggplot(aes(longitude, latitude, color = pr)) +
  borders(database = "world", regions = "Kenya") +
  geom_point() +
  scale_color_gradient2(low = "blue", midpoint = 0.5, high = "red",
                        labels = percent_format()) +
  coord_map() +
  theme_void()

```

```{r}
# map by decade!
kenya_pr %>% 
  mutate(decade = 10 * year_start %/% 10) %>% 
  arrange(pr) %>% 
  ggplot(aes(longitude, latitude, color = pr)) +
  borders(database = "world", regions = "Kenya") + 
  geom_point() + 
  scale_color_gradient2(low = "blue", mid = "orange", high = "red", midpoint = 0.5) +
  facet_wrap(~ decade) +
  theme_void() +
  coord_map() +
  labs(color = "Prevalence")

```



Aggregated across countries


**Malaria inicdents**

```{r}
malaria_inc <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-11-13/malaria_inc.csv")
malaria_inc

malaria_inc_processed <- malaria_inc %>% 
  set_names(c("country", "code", "year", "incidence")) %>% 
  mutate(incidence = incidence / 1000)  # percent
```

```{r}
malaria_inc_processed %>% 
  filter(country %in% sample(unique(country), 6)) %>% # sample 6 countries randomly => get a feeling for the data set
  ggplot(aes(year, incidence, color = country)) +
  geom_line() +
  scale_y_continuous(labels = percent_format())

```


Looking at 2015 levels and the changes form 2000 to 2015

```{r}
malaria_inc_processed %>% 
  group_by(country) %>% 
  summarize(current = last(incidence),
            change = last(incidence) / first(incidence)) %>% 
  arrange(desc(change))


# a bit more detailed approach

malaria_spread <- malaria_inc_processed %>% 
                  mutate(year = paste0("Y", year)) %>% 
                  spread(year, incidence)
                  

malaria_spread %>% 
  filter(country != "Turkey",
         !is.na(code)) %>% 
  mutate(current = Y2015,
         change = Y2015 - Y2000) %>% # difference instead of ratio
  arrange(change) %>% 
  ggplot(aes(current, change)) +
  geom_point() +
  geom_text(aes(label = code))

```


Map the data!
```{r}
malaria_spread %>% 
  anti_join(map_data("world"), by = c(country = "region")) # good for merging!
```

Merging & mapping (malaria inicidence over time)
```{r}

world <- map_data("world") %>% 
  filter(region != "Antarctica")

malaria_spread %>% 
  inner_join(maps::iso3166 %>% 
               select(a3, mapname), by = c(code = "a3")) %>%  # we do not need every column...
  right_join(world, by = c(mapname = "region")) %>% 
  
  ggplot(aes(long, lat, group = group, fill = Y2015)) +
  geom_polygon() +
  scale_fill_gradient2(low = "blue", mid = "orange", high = "red", midpoint = 0.25,
                       label = percent_format()) +
  theme_void() +
  coord_fixed()


malaria_inc_processed %>% 
  filter(incidence < 1) %>% 
  inner_join(maps::iso3166 %>% 
               select(a3, mapname), by = c(code = "a3")) %>%  # we do not need every column...
  left_join(world, by = c(mapname = "region")) %>% 
  
  ggplot(aes(long, lat, group = group, fill = incidence)) +
  geom_polygon() +
  scale_fill_gradient2(low = "blue", mid = "orange", high = "red", midpoint = 0.35,
                       label = percent_format()) +
  theme_void() +
  coord_fixed() +
  facet_wrap(~ year) +
  labs(title = "Malaria incidence over time")
```


**Malaria deaths**

```{r}
malaria_deaths <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-11-13/malaria_deaths.csv")

malaria_deaths_processed <- malaria_deaths %>% 
  set_names(c("country", "code", "year", "deaths")) # per 100'000 (not %)

malaria_deaths_processed

```


```{r}
malaria_deaths_processed %>% 
  filter(country %in% sample(unique(country), 6)) %>% # sample 6 countries randomly => get a feeling for the data set
  ggplot(aes(year, deaths, color = country)) +
  geom_line() +
  labs(y = "Deaths per 100'000")

```


```{r}
world

malaria_deaths_processed %>% 
  filter(year %% 10 == 0) %>% 
  inner_join(maps::iso3166 %>% 
               select(a3, mapname), by = c(code = "a3")) %>% 
  inner_join(world, by = c(mapname = "region")) %>% 
  ggplot(aes(long, lat, fill = deaths, group = group)) +
  geom_polygon() +
  scale_fill_gradient2(low = "blue", mid = "orange", high = "red", midpoint = 100) +
  coord_fixed() +
  facet_wrap(~ year) +
  theme_void() +
  labs(title = "Malaria deaths over time around the world")

# china is missing!
```

```{r}
malaria_country_data <- malaria_deaths_processed %>% 
  inner_join(maps::iso3166 %>% 
               select(a3, mapname), by = c(code = "a3")) %>% 
  mutate(mapname = str_remove(mapname, "\\(.*"))

malaria_country_data %>% 
  filter(mapname == "China")

# worked


world %>% 
  inner_join(malaria_country_data, by = c(region = "mapname")) %>% 
  filter(year %% 10 == 0) %>% 
  ggplot(aes(long, lat, fill = deaths, group = group)) +
  geom_polygon() +
  scale_fill_gradient2(low = "blue", mid = "orange", high = "red", midpoint = 100) +
  coord_fixed() +
  facet_wrap(~ year) +
  theme_void() +
  labs(title = "Malaria deaths over time around the world")
```

















