---
title: "TidyTemplate"
date: 2021-06-16
output: html_output
editor_options: 
  chunk_output_type: console
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
library(lubridate)
theme_set(theme_light())
```

# Load the weekly Data

Dowload the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2021-03-23")

unvotes <- tt$unvotes %>% 
  mutate(vote_number = match(vote, c("no", "abstain", "yes")) -2) %>% 
  left_join(tt$roll_calls %>% 
              select(rcid, date, amend), by = "rcid")
  
unvotes

unvotes %>% 
  count(vote, sort = TRUE) 

unvotes %>% 
  count(country, sort = TRUE)
```


Create a function
```{r}
summarize_votes <- function(tbl, min_votes = 10) {
  tbl %>% 
    summarize(n_votes = n(),
              n_yes = sum(vote == "yes"),
              pct_yes = n_yes / n_votes,
              .groups = "drop") %>% 
    filter(n_votes >= min_votes) %>% 
    arrange(desc(pct_yes))
}
```


```{r}
by_country <- unvotes %>% 
  group_by(country, country_code) %>% 
  summarize_votes()



by_country %>% 
  slice(c(1:10, (n() - 10):n())) %>% 
  mutate(country = fct_reorder(country, pct_yes)) %>%
  ggplot(aes(pct_yes, country)) +
  geom_point(aes(size = n_votes)) +
  scale_x_continuous(labels = percent) + 
  labs(x = "% of yes votes in UN",
       title = "What countries voted yes the least/the most?")
```


```{r}
by_year <- unvotes %>% 
  group_by(year = year(date)) %>% 
  summarize_votes()

by_year

by_year %>% 
  ggplot(aes(year ,pct_yes)) + 
  geom_line() +
  expand_limits(y = 0)

by_country_year <- unvotes %>% 
  group_by(year = year(date), country, country_code) %>% 
  summarize_votes()

by_country_year %>% 
  filter(country %in% c("United States", "Canada", "Mali", "Israel", "Germany", "France")) %>% 
  mutate(country = fct_reorder(country, pct_yes)) %>% 
  ggplot(aes(year, pct_yes, color = country)) + 
  geom_line() + 
  scale_color_discrete(guide = guide_legend(reverse = TRUE)) +
  expand_limits(y = 0) +
  labs(y = "% yes votes over time")


by_country_year %>% 
  filter(country %in% c("United States", "Canada", "Mali", "Israel", "Germany", "France")) %>% 
  mutate(country = fct_reorder(country, pct_yes)) %>% 
  ggplot(aes(year, pct_yes, color = country)) + 
  geom_line() + 
  scale_y_continuous(labels = percent) +
  expand_limits(y = 0) +
  facet_wrap(~ country) +
  theme(legend.position = "none") +
  labs(y = "% yes votes over time")


# adding an overall
by_country_year %>%
  filter(country %in% c("United States", "Canada", "Mali", "Israel", "Germany", "France")) %>% 
  mutate(country = fct_reorder(country, pct_yes)) %>% 
  ggplot(aes(year, pct_yes)) +
  geom_line(data = by_year, size = 1.5) +
  geom_line(aes(color = country)) +
  scale_y_continuous(labels = percent) +
  scale_color_discrete(guide = guide_legend(reverse = TRUE)) +
  expand_limits(y = 0) +
  labs(y = "% yes votes over time")
  
```


World map

```{r}
library(ggthemes)
library(fuzzyjoin)

world_data <- map_data("world") %>% 
  as_tibble() %>% 
  regex_left_join(maps::iso3166 %>% 
                    select(mapname, country_code = a2), 
                  c(region = "mapname")) %>% 
  filter(region != "Antarctica")

world_data


world_data %>% 
  left_join(by_country, by = "country_code") %>% 
  ggplot(aes(long, lat, group = group, fill = pct_yes)) +
  geom_polygon() +
  theme_map() + 
  scale_fill_gradient2(low = "red",
                       high = "blue",
                       midpoint = 0.6,
                       labels = percent) +
  labs(fill = "% yes votes")
```


By continent

```{r}
library(countrycode)

# create another function
plot_by <- function(tbl, category) {
  tbl %>% 
    filter(!is.na({{ category }})) %>% 
    mutate(category = fct_reorder({{ category }}, pct_yes)) %>% 
    ggplot(aes(year, pct_yes)) +
    geom_line(aes(color = category)) +
    scale_y_continuous(labels = percent) +
    scale_color_discrete(guide = guide_legend(reverse = TRUE),
                       name = "Continent") +
    expand_limits(y = 0) +
    labs(y = "% Yes Votes Over Time",
         x  = "Year")
  }


unvotes %>% 
  group_by(continent = countrycode(country_code, "iso2c", "continent"),
           year = year(date)) %>% 
  summarize_votes() %>% 
  plot_by(continent)


```


```{r}
library(WDI)

country_incomes <- WDI(indicator = c(gdp_per_capita = "NY.GDP.PCAP.PP.KD",
                                     pop = "SP.POP.TOTL"), 
                       start = 2000, end = 2019,
                       extra = TRUE) %>% 
  as_tibble() %>% 
  select(country_code = iso2c, year,  income, gdp_per_capita, pop) %>% 
  filter(!is.na(income)) %>% 
  mutate(income = fct_relevel(income, "Low income", "Lower middle income", "Upper middle income"))

country_incomes


unvotes %>% 
  inner_join(country_incomes, by = "country_code") %>% 
  group_by(income,
           year = year(date)) %>% 
  summarize_votes() %>% 
  plot_by(income)

```


Correlation between countries

```{r}
unvotes %>% 
  filter(country %in% c("United States", "Canada")) %>% 
  select(rcid, country, vote_number) %>% 
  spread(country, vote_number, fill = 0) %>% 
  summarize(correlation = cor(Canada, `United States`))


# do it all at the same time for one specific country!
library(widyr)
country_correlations <- unvotes %>% 
  as_tibble() %>% 
  filter(country != "Zanzibar") %>% 
  pairwise_cor(country, rcid, vote_number, sort = TRUE) 



country_correlations %>% 
  filter(item1 == "Russia") %>%                                        # choose whatever country you like!
  slice(c(1:20, n() - 0:19)) %>% 
  mutate(country = fct_reorder(item2, correlation)) %>% 
  ggplot(aes(correlation, country)) + 
  geom_errorbarh(height = 0, aes(xmin = correlation, xmax = 0)) +
  geom_point() +
  labs(x = "How much this country's voting is correlatied with Russia",
       y = "")
  
country_correlations %>% 
  filter(item1 == "United States") %>%                                        # choose whatever country you like!
  slice(c(1:20, n() - 0:19)) %>% 
  mutate(country = fct_reorder(item2, correlation)) %>% 
  ggplot(aes(correlation, country)) + 
  geom_errorbarh(height = 0, aes(xmin = correlation, xmax = 0)) +
  geom_point() +
  labs(x = "How much this country's voting is correlatied with the US",
       y = "")

country_correlations %>% 
  filter(item1 == "North Korea") %>%                                        # choose whatever country you like!
  slice(c(1:20, n() - 0:19)) %>% 
  mutate(country = fct_reorder(item2, correlation)) %>% 
  ggplot(aes(correlation, country)) + 
  geom_errorbarh(height = 0, aes(xmin = correlation, xmax = 0)) +
  geom_point() +
  labs(x = "How much this country's voting is correlatied with North Korea",
       y = "")
```

Which countries agree/disagree the most?

```{r}
country_correlations %>% 
  group_by(country = item1) %>% 
  summarize(median_correlation = median(correlation),
            avg_correlation = mean(correlation)) %>% 
  arrange(desc(median_correlation))
```

Text analysis

```{r}
library(tidytext)

rc_words <- tt$roll_calls %>% 
  filter(!is.na(short)) %>% 
  unnest_tokens(word, short) %>%
  anti_join(stop_words, by = "word") %>% 
  select(rcid, word)


unvotes %>% 
  inner_join(rc_words, by = "rcid") %>% 
  filter(country == "United States") %>% 
  group_by(word) %>% 
  summarize_votes(min_votes = 100) %>% 
  mutate(word = fct_reorder(word, pct_yes)) %>% 
  ggplot(aes(pct_yes, word)) +
  geom_point(aes(size = n_votes)) +
  expand_limits(x = 0) +
  scale_x_continuous(labels = percent) +
  labs(x = "% yes")
```





































