---
title: "World Freedom Index"
author: "Luca"
date: "28/02/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r}
library(tidyverse)
library(tidytuesdayR)
library(ggrepel)
library(scales)
library(broom)
library(glue)
library(countrycode)


theme_set(theme_light())
```


Download the data

```{r}
freedom <- tt_load("2022-02-22")$freedom %>% 
  janitor::clean_names() %>% 
  rename(civil_liberties = cl,
         political_rights = pr) %>% 
  # fix with the countrycode-package!
  mutate(country_code = countrycode(country, "country.name", "iso2c"))

freedom

freedom %>% 
  count(year) %>% 
  arrange(year)
```

from 1995 to 2020
about 193 countries (some missing data in earlier years)


Civil liberty

```{r}
freedom %>% 
  ggplot(aes(civil_liberties)) +
  geom_histogram()


freedom %>% 
  filter(year == 2020) %>% 
  arrange(desc(civil_liberties))

# scales from 1 to 7 (1 = best, 7 = worst)

```


Look just at 2020

```{r}
# create function

summarise_freedom <- function(tbl) {
  tbl %>% 
    summarize(n_countries = n(),
              avg_civil_liberties = mean(civil_liberties),
              avg_political_rights = mean(political_rights),
              pct_free = mean(status == "F"),
              .groups = "drop") %>% 
    arrange(desc(n_countries))
}
```


Correlation between region and freedom indices

```{r}
by_region <- freedom %>% 
  filter(year == 2020) %>% 
  group_by(region_name) %>% 
  summarise_freedom()


by_region %>% 
  ggplot(aes(avg_civil_liberties, avg_political_rights)) + 
  geom_point(aes(size = n_countries),
             alpha = 0.25) +
  geom_abline(color = "red") + 
  geom_text(aes(label = glue("{region_name}: {round(pct_free, 1)}% countries are free")),
            color = "black",
            vjust = 0,
            hjust = 1.05) +
  expand_limits(x = 0,
                y = 0,
                size = 0) +
  scale_x_continuous(breaks = seq(0, 5, 1)) +
  scale_y_continuous(breaks = seq(0, 5, 1)) + 
  labs(title = "Freedom Indices by Region",
       x = "Civil Liberties",
       y = "Political Rights",
       size = "Number of Countries")
```


Correlation between countries and freedom indices

```{r}
# is the mean the right value?
freedom %>% 
  filter(year == 2020) %>% 
  ggplot(aes(civil_liberties, political_rights)) +
  geom_jitter(width = 0.2, 
              height = 0.2) +
  geom_abline(color = "red") + 
  scale_x_continuous(breaks = seq(0, 7, 1)) + 
  scale_y_continuous(breaks = seq(0, 7, 1)) + 
  expand_limits(x = 0, 
                y = 0) + 
  labs(title = "Freedom Indices by Country",
       x = "Civil Liberties",
       y = "Political Rights")

# look at the distribution as well!
freedom %>% 
  filter(year == 2020) %>% 
  pivot_longer(cols = c(civil_liberties, political_rights), 
               names_to = "metric",
               values_to = "value") %>% 
  mutate(metric = str_to_title(str_replace_all(metric, "_", " ")),
         region_name = fct_reorder(region_name, value)) %>% 
  count(region_name, metric, value) %>% 
  ggplot(aes(value, n)) + 
  geom_col() +
  scale_x_continuous(breaks = seq(1, 7, 1))  +
  facet_grid(region_name ~ metric) +
  labs(title = "Distribution of Freedom Index Rating per Region",
       x = "Freedom Index Rating",
       y = "Number of Countries")
  
```


Over time

```{r}
freedom_long <- freedom %>% 
  pivot_longer(cols = c(civil_liberties, political_rights),
               names_to  = "metric",
               values_to = "value") %>% 
  mutate(metric = str_to_title(str_replace_all(metric, "_", " ")),
         region_name = fct_reorder(region_name, value))


# create second data set for the overall trend!
overall <- freedom_long %>% 
  group_by(year, metric) %>% 
  summarise(avg_rating = mean(value))

freedom_long %>% 
  group_by(year, region_name, metric) %>% 
  summarise(avg_rating = mean(value)) %>% 
  ggplot(aes(year, avg_rating)) +
  geom_line(aes(color = region_name)) +
  geom_line(data = overall,
            aes(year, avg_rating),
            size = 2) + 
  facet_wrap(~ metric) + 
  
  # more intuive to understand!
  scale_y_reverse(breaks = seq(1, 5, 1)) +
  expand_limits(y = 1) + 
  
  # change order of the legend!
  #scale_color_discrete(guide = guide_legend(reverse = TRUE)) +
  
  labs(title = "Word Freedom Index Rating over Time",
       subtitle = "Overall trend is represented in black",
       x = "Year",
       y = "Word Freedom Index Rating",
       color = "Region")

# other trick video (24:00)
#freedom_long %>% 
#  mutate(region_name = "Overall") %>% 
#  bind_rows(freedom_long) %>% 

```





World bank data

```{r}
library(WDI)
gdp_percap <- WDI(indicator = "NY.GDP.PCAP.CD",
                  extra = TRUE,
                  start = 1995,
                  end = 2020) %>% 
  as_tibble()
```

Merge with the freedom data
```{r}
freedom_joined <- freedom %>% 
  inner_join(gdp_percap,
             by = c(country_code = "iso2c", "year"),
             suffix = c("", "_wdi"))

# 2020
freedom_joined %>% 
  filter(year == 2020,
         income != "Not classified") %>% 
  mutate(income = fct_relevel(income,
                              c("Low income", "Lower middle income", "Upper middle income", "High income"))) %>% 
  group_by(income) %>% 
  summarise_freedom() %>% 
  arrange(income)


# by year
freedom_joined <- freedom_long %>% 
  inner_join(gdp_percap, by = c(country_code = "iso2c", "year"),
             suffix = c("", "_wdi")) %>% 
  mutate(income = fct_relevel(income,
                              c("Low income", "Lower middle income", "Upper middle income", "High income")))

freedom_joined %>% 
  filter(income != "Not classified") %>% 
  group_by(metric, income, year) %>% 
  summarise(avg_rating = mean(value)) %>% 
  ggplot(aes(year, avg_rating)) +
  geom_line(aes(color = income)) + 
  geom_line(data = overall,
            aes(year, avg_rating),
            size = 2) +
  scale_y_reverse(breaks = seq(1, 5, 1)) +
  expand_limits(y = 1) + 
  facet_wrap(~ metric) + 
  labs(title = "World Freedom Index Rating over Time",
       subtitle = "Overall trend is represented in black",
       x = "Year",
       y = "World Freedom Index Rating")

```


Run a regression
Replication problem with time series data
=> just look at one year...
=> if we would just take into the model, we would be to confident with our predictions
=> one would have to run a propre times series regression
```{r}
# freedom regions
civil_liberties_2020 <- freedom_joined %>% 
  filter(metric == "Civil Liberties",
         year == 2020,
         !is.na(NY.GDP.PCAP.CD))


lin_mod <- civil_liberties_2020 %>% 
  # Europe = base line
  lm(value ~ region_name + log2(NY.GDP.PCAP.CD), data = .)

summary(lin_mod)

# WDI regions
lin_mod_2 <- civil_liberties_2020 %>% 
  lm(value ~ region + log2(NY.GDP.PCAP.CD), data = .)

summary(lin_mod_2)  
```


What countries in each group are the greatest outliers compared to the avg trend?
Where did our model fail to predict the actual value?
```{r}
lin_mod %>% 
  augment(data = civil_liberties_2020) %>% 
  select(country, region_name, NY.GDP.PCAP.CD, income, value, .fitted, .resid) %>% 
  arrange(desc(abs(.resid)))

# plot
lin_mod %>% 
  augment(data = civil_liberties_2020) %>% 
  select(country, region_name, NY.GDP.PCAP.CD, income, value, "fitted" = .fitted, "residuals" = .resid) %>% 
  arrange(desc(abs(residuals))) %>% 
  head(20) %>% 
  ggplot(aes(fitted, value)) +
  geom_point() +
  geom_abline(color = "red") +
  scale_x_continuous(breaks = seq(1, 7, 1)) + 
  scale_y_continuous(breaks = seq(1, 7, 1)) +
  expand_limits(x = 1, 
                y = 1) +
  geom_text_repel(aes(label = country)) +
  labs(title = "What are the largest outliers?",
       x = "Fitted Value based on Region and Income",
       y = "Actual Freedom Index")
```


Create a map for 2020
```{r}
library(maps)
library(fuzzyjoin)

world <- map_data("world") %>% 
  as_tibble()

freedom_2020 <- freedom_joined %>% 
  filter(year == 2020)

world_map_freedom_2020 <- world %>% 
  regex_left_join(maps::iso3166, c(region = "mapname")) %>% 
  left_join(freedom_2020 %>% select(-region), 
            by = c(a2 = "country_code")) %>% 
  filter(region != "Antarctica")


world_map_freedom_2020 %>% 
  filter(metric == "Civil Liberties") %>% 
  ggplot(aes(long, lat, group = group)) + 
  geom_polygon(aes(fill = value)) + 
  
  # coord_map fix!
  coord_map(xlim = c(-180, 180)) + 
  
  scale_fill_gradient2(low = "blue",
                       midpoint = 3.5,
                       high = "red",
                       guide = guide_legend(reverse = TRUE)) +
  ggthemes::theme_map() + 
  labs(title = "World Freedom Index: Civil Liberties by Country",
       subtitle = "In 2020",
       fill = "Civil Liberties")
```


create animated map with gganimate
```{r}
library(gganimate)

world_map_freedom <- world %>% 
  regex_left_join(maps::iso3166, c(region = "mapname")) %>% 
  left_join(freedom_joined %>% select(-region),
            by = c(a2 = "country_code")) %>% 
  filter(region != "Antartica")


world_map_freedom %>% 
  filter(metric == "Civil Liberties") %>% 
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(aes(fill = value)) + 
  coord_fixed(1.25) + 
  scale_fill_gradient2(low = "blue",
                       midpoint = 3.5,
                       high = "red",
                       guide = guide_legend(reverse = TRUE)) + 
  ggthemes::theme_map() +
  transition_manual(year) + 
  labs(title  = "World Freedom Index: Civil Liberties by Country",
       subtitle = "In { current_frame }",
       fill = "Civil Liberties")

```


Write a function
```{r}

detach("package:gganimate")

for (i in 1995:2020) {
  plot <- world_map_freedom %>% 
  filter(metric == "Civil Liberties",
         year == i) %>% 
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(aes(fill = value)) + 
  coord_fixed(1.25) + 
  scale_fill_gradient2(low = "blue",
                       midpoint = 3.5,
                       high = "red",
                       guide = guide_legend(reverse = TRUE)) + 
  ggthemes::theme_map() +
  labs(title  = "World Freedom Index: Civil Liberties by Country",
       subtitle = glue("In {i}"),
       fill = "Civil Liberties")
  
  # store as separate plot
  assign(paste("freedom_plot", i, sep = "_"), plot)
  
  # remve plot
  rm(plot)
}
```

