---
title: "Beer Production - Bootstrap Resampling"
author: "Luca"
date: '2022-05-07'
output: html_document
editor_options: 
  chunk_output_type: console
---

## Load packages and get the data

### Data Dictionary

variable                    class               description

data_type                 	character 	        Pounds of Material - this is a sub-table from beer_taxed
material_type             	character         	Grain product, Totals, Non-Grain Product (basically hops vs grains)
year 	                      double 	            Year
month                     	integer            	Month
type                        character         	Actual line-item from material type
month_current             	double 	            Current number of barrels for this year/month
month_prior_year          	double 	            Prior year number of barrels for same month
ytd_current               	double            	Cumulative year to date of current year
ytd_prior_year            	double 	            Cumulative year to date for prior year

```{r}
library(tidyverse)
library(tidymodels)

knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE,
                      echo  = TRUE,
                      dpi = 100,
                      fig.width = 8,
                      fig.height = 5)

theme_set(theme_light())

brewing_materials_raw <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/brewing_materials.csv")

```

A bit of EDA

```{r}
brewing_materials_raw

# do some counting
for(var in names(brewing_materials_raw)) {
  
  print(
    brewing_materials_raw %>% 
      count(.data[[var]], sort = TRUE)
  )
  
}


# what is used the most?
brewing_materials_raw %>% 
  count(type, wt = month_current, sort = TRUE)
```



```{r}
brewing_filtered <- brewing_materials_raw %>%
  filter(type %in% c("Malt and malt products",
                     "Sugar and syrups",
                     "Hops (dry)")) %>% 
  filter(year < 2016,
         month != 12) %>% 
  mutate(date = str_c(year, "-", month, "-01"),
         date = lubridate::ymd(date))


brewing_filtered %>% 
  ggplot(aes(date, month_current, color = type))  +
  geom_point()
```


```{r}
brewing_materials <- brewing_filtered %>% 
  select(date, type, month_current) %>% 
  pivot_wider(names_from = type,
              values_from = month_current) %>% 
  janitor::clean_names()


brewing_materials %>% 
  ggplot(aes(malt_and_malt_products, sugar_and_syrups)) +
  geom_point() + 
  geom_smooth(method = "lm")
```



## Fit a "simple" lm

```{r}
# we do not include the intercept. Assumption: "no malt no sugar needed"
beer_fit <- lm(sugar_and_syrups ~ 0 + malt_and_malt_products, data = brewing_materials)

summary(beer_fit)
```


## Bootstrap resampling

Lets fit the model on more data to get more robust estimations

```{r}
# create 1000 boostprap resamples
set.seed(1234)
beer_boot <- bootstraps(brewing_materials,
                        times = 1000,
                        # apparent needs to be true to calculate the conf.int. (see below!)
                        apparent = TRUE)

beer_boot
```



```{r}
# fit the model on these 1000 samples
# map takes an object (dataframe, list, vector etc.) and maps a function to it!

beer_models <- beer_boot %>% 
  
  # first lets maps over all splits and use the lm-function (same as above)
  mutate(model = map(splits, ~lm(sugar_and_syrups ~ 0 + malt_and_malt_products, 
                                 data = .)),
         
         # lets map over each model and tidy it => get the model informations!
         coef_info = map(model, tidy))

# now we have fittet the model to all of the 1000 samples!
# this is allowes us to get much more robust estimates!
```


## Evaluate the results

```{r}
# take a look at the coefficients
beer_coefs <- beer_models %>% 
  unnest(coef_info)

# lets take a look at the distribution of the estimates!
beer_coefs %>% 
  ggplot(aes(estimate)) + 
  geom_histogram(alpha = 0.8)
```


```{r}
# we can even calculate the conf.intervals! 
conf_int <- int_pctl(.data = beer_models,
                     statistics = coef_info,
                     alpha = 0.05)

# Interpretation: beer producer need about 5 barrles malt for every one barrel of sugar

# add the conf.int to the plot 
beer_coefs %>% 
  ggplot(aes(estimate)) +
  geom_histogram(alpha = 0.8) + 
  geom_vline(aes(xintercept = .lower), 
             data = conf_int,
             color = "red",
             size = 1.5) + 
  geom_vline(aes(xintercept = .upper),
             data = conf_int,
             color = "red",
             size = 1.5)
```


```{r}
# lets visualize the residuals for each model
# use map with augment to get the true values (sugar_and_syrups) and the fitted values (.fitted)
beer_aug <- beer_models %>% 
  mutate(augmented = map(model, augment)) %>% 
  unnest(augmented)

# plot
# the points are the real values
# the set of lines are the predicted values of each of the 1000 models!
# So you can see the varation between the models
beer_aug %>% 
  ggplot(aes(malt_and_malt_products, sugar_and_syrups))  +
  geom_line(aes(y = .fitted, group = id),
            alpha = 0.1,
            color = "red") +
  geom_point()
```

