---
title: "Tidy Tuesday Horror Movie Ratings"
author: "Luca"
date: "29/01/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
theme_set(theme_light())

horror_movies_raw <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-22/horror_movies.csv")
```


Most of the movies are since 2012

```{r}
horror_movies <- horror_movies_raw %>% 
  arrange(desc(review_rating)) %>% 
  extract(title, "year", "\\((\\d\\d\\d\\d)\\)$",
          remove = FALSE,
          convert = TRUE) %>% 
  mutate(budget2 = parse_number(budget)) %>% 
  # remove "Directed by (...)" and "With (...)
  separate(plot, c("director", "cast_sentence", "plot"),
           extra = "merge",
           sep = "\\. ",
           fill = "right") %>% 
  distinct(title, .keep_all = TRUE)


# only NAs could not be parsed =>  good
horror_movies %>% 
  filter(is.na(budget2)) %>% 
  count(budget, sort = TRUE)


horror_movies <- horror_movies %>% 
  select(-budget) %>% 
  rename("budget" = budget2)
```


"Data quality" - Proportion of NAs

```{r}
horror_movies %>% 
  summarise(across(everything(), ~mean(is.na(.)))) %>% 
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Proportion_NA") %>% 
  mutate(Percent_NA = Proportion_NA * 100)
```


```{r}
horror_movies %>% 
  count(genres, sort = TRUE)

horror_movies %>% 
  count(language, sort = TRUE)


horror_movies %>% 
  ggplot(aes(budget)) +
  geom_histogram() + 
  scale_x_log10(labels = scales::dollar_format())
  
```


Do higher budget movies end up higher rated?

```{r}
horror_movies %>% 
  ggplot(aes(budget, review_rating)) +
  scale_x_log10(labels = scales::dollar_format()) +
  geom_point() + 
  geom_smooth(method = "lm")


cor(x = log10(horror_movies$budget),
    y = horror_movies$review_rating,
    use = "complete.obs")

```

No realtionship between budget and rating...

How about movie rating and review?

```{r}
horror_movies %>% 
  mutate(movie_rating = fct_lump(movie_rating, 5),
         movie_rating = fct_reorder(movie_rating, review_rating,
                                    na.rm = TRUE)) %>%
  ggplot(aes(movie_rating, review_rating)) + 
  geom_boxplot() + 
  coord_flip()

# is there a relationship? => anova (analyse the variance)!
horror_movies %>%
  filter(!is.na(movie_rating)) %>% 
  mutate(movie_rating = fct_lump(movie_rating, 5)) %>% 
  lm(review_rating ~ movie_rating, data = .) %>% 
  anova()
  
# F-test is significant => there is more variation than we would expect by chance...
```


```{r}
horror_movies %>% 
  # movies in more than one genre get copied (Horror as baseline, because every movie is in it)
  separate_rows(genres, sep = "\\| ") %>% 
  mutate(genre = fct_lump(genres, 5),
         genre = fct_reorder(genre, review_rating,
                             na.rm = TRUE)) %>% 
  ggplot(aes(genre, review_rating)) + 
  geom_boxplot() +
  coord_flip()

```


```{r}
library(tidytext)

# every token (= here set to a word) in plot gets his own row
horror_movies_unnested <- horror_movies %>% 
  unnest_tokens(word, plot) %>% 
  # drops all observations in x that have a match in y
  anti_join(stop_words,
            by = "word") %>% 
  filter(!is.na(word))


horror_movies_unnested %>% 
  filter(!is.na(review_rating)) %>% 
  group_by(word) %>% 
  summarize(movies = n(),
            avg_rating = mean(review_rating)) %>% 
  # words that appear in at least 100 movie plot descriptions
  arrange(desc(movies)) %>% 
  filter(movies >= 100) %>% 
  mutate(word = fct_reorder(word, avg_rating)) %>% 
  ggplot(aes(avg_rating, word)) + 
  geom_point()
```

### Lasso regression for predicting review rating based in words

```{r}
library(glmnet)
library(Matrix)

movie_word_matrix <- horror_movies_unnested %>% 
  filter(!is.na(review_rating)) %>% 
  # we use only words, that appear at least 20 times
  add_count(word) %>% 
  filter(n >= 20) %>%
  count(title, word) %>% 
  cast_sparse(row = title, column = word, value = n)

# 2945 movies, 460 words  
dim(movie_word_matrix)
```


Which words in the plot predict a higher or lower rating?

```{r}
# "create" y-variable
rating <- horror_movies$review_rating[match(rownames(movie_word_matrix), horror_movies$title)]

# almost normally distributed
qplot(rating)

# cross validation glmnet
lasso_model <- cv.glmnet(movie_word_matrix, rating)

# broom-package
library(broom)
tidy(lasso_model$glmnet.fit) %>% 
  View()

# NOTE: we use over 450 predictors to predict the rating
# it is very likely that we overfit the model
# lambda is a penalty-term => "the smaller lambda the further away are the terms from 0"
#                             (a high lambda sets predictors to 0's...)
tidy(lasso_model$glmnet.fit) %>% 
  filter(term %in% c("house", "friends", "life", "family", "horror", "evil", "mother")) %>% 
  ggplot(aes(lambda, estimate, color = term)) + 
  geom_line() + 
  scale_x_log10() + 
  geom_hline(yintercept = 0, lty = 2)
# how to choose lambda?!
```

Choosing lambda

```{r}
# take a look at the Mean-Squared Error => gets very big => overfitting of the model!
plot(lasso_model)

lasso_model$lambda.min

tidy(lasso_model$glmnet.fit) %>% 
  filter(lambda == lasso_model$lambda.min,
         term != "(Intercept)") %>% 
  mutate(term = fct_reorder(term, estimate)) %>% 
  ggplot(aes(term, estimate)) +
  geom_col() +
  coord_flip()
```



Model with director, cast, genre, rating, plot, words, language, release country

```{r}
features <- horror_movies %>% 
  filter(!is.na(review_rating)) %>% 
  select(title, genres, director, cast, movie_rating, language, release_country) %>% 
  mutate(director = str_remove(director, "Directed by ")) %>% 
  pivot_longer(cols = -title, names_to = "type", values_to = "value")  %>%
  filter(!is.na(value)) %>% 
  separate_rows(value, sep = "\\| ?") %>% 
  unite(feature, type, value, sep = ": ") %>% 
  mutate(n = 1)

movie_feature_matrix <- horror_movies_unnested %>% 
  filter(!is.na(review_rating)) %>% 
  count(title, feature = paste0("word: ", word)) %>% 
  bind_rows(features) %>% 
  add_count(feature, name = "counter") %>% 
  filter(counter >= 10) %>% 
  cast_sparse(title, feature)


# 3051 movies, 1142 predictors
dim(movie_feature_matrix)

```


```{r}
rating <- horror_movies$review_rating[match(rownames(movie_feature_matrix), horror_movies$title)]

feature_lasso_model <- cv.glmnet(movie_feature_matrix, rating)
```

```{r}
plot(feature_lasso_model)

tidy(feature_lasso_model$glmnet.fit) %>% 
  filter(lambda == feature_lasso_model$lambda.1se,
         term != "(Intercept)") %>% 
  separate(term, "type", sep = ":", remove = FALSE) %>%
  mutate(type = str_to_title(type),
         type = str_replace_all(type, "_", " "),
         term = str_replace_all(term, "^[^:]+", ""),
         term = str_replace_all(term, ": ", ""),
         term = str_to_title(term),
         term = fct_reorder(term, estimate)) %>% 
  ggplot(aes(term, estimate, fill = type)) + 
  geom_col() +
  coord_flip() +
  labs(x = "",
       y = "Coefficient for predicting horror movie rating",
       fill = "Type",
       title = "What affects a horror movie rating?",
       subtitle = "Based on a lasso regression to predict IMDb ratings of ~3000 movies")

```

