---
title: "One-Way ANOVA"
author: "Luca"
date: '2022-06-03'
output: html_document
---

Check out the full tutorial by David Caughlin [here](https://www.youtube.com/watch?v=yoPGwvUzjgQ&list=PLKkRkURCtPjCJOZHskCoyJCPb8wMDs2CW&index=23)


### load packages

```{r}
library(tidyverse)
library(psych)
library(lessR)
library(effsize)

theme_set(theme_light())
```

### load data

```{r}
surveydata <- read_csv("https://raw.githubusercontent.com/davidcaughlin/R-Tutorial-Data-Files/master/TrainingEvaluation_ThreeGroupPost.csv")

surveydata
```


## Introdcution

ANOVA (Analysis of Variance)
one-way because we only have one factor (Condition) as a predictor variable.
NOTE: but this predictor has 3 levels => need to use ANOVA istead of t-test...
Question: what is the best Conditon (No vs. Old vs. New) => are the means between these groups statistically different from one another?


### statistical assumptions

```{r}
### visualize statistical assumptions

# are the distributions of the outcome variable for each condition normally distributed?
# equal variance across the conditions?
Plot(PostTest,
     by1 = Condition,
     data = surveydata)

# looks good



### statistical tests

# normality assumption
surveydata %>% 
  split(.$Condition) %>% 
  map(~ shapiro.test(.$PostTest))
# high p-values (H0 cannot be rejected) => normal distribution for all Conditions!



# equality of variances
car::leveneTest(PostTest ~ Condition,
                data = surveydata)
# high p-values (H0 cannot be rejected) => equality of variances!
```



### run one-way ANOVA

```{r}
ANOVA(PostTest ~ Condition,
      data = surveydata)


### interpretation of the plots

# 1. 95% family-wise confidence level
# if 0 is included in the interval => not statistically significant (the difference between Old and New is not statistically significant)
# if 0 is not included in the interval => statistically significant (the differences between No and New AND Old and No is statistically significant)

# 2. scatterplot with the distribution and the mean values for each group (Condition)


### interpretation of console output

# ANOVA
# significant F-test (high F-values and very low p-value) => H0 (no differences in means) can be rejected!
# => the 3 means are statistically different from each other!
# NOTE: this is an omnibus test. So it does not tell us which means are different! It just states that statistically significant differences exist.


# Association and Effect Size
# R Squared: 0.220 (22 % of the variance of the output variable can be explained by Condition)
# R Sq Adjusted and Omega Squared => measures on the population level...
# the size of these measurements indicate a medium to large effect size

# Cohen's f: 0.494 (indicates a large effect size)

# NOTE: these measurements are also on an omnibus level...



# but which pairs are significantly different from one another?
#   TUKEY MULTIPLE COMPARISONS OF MEANS 
# 
# Family-wise Confidence Level: 0.95 
# ----------------------------------- 
#             diff    lwr   upr p adj 
#    No-New -10.00 -15.48 -4.52  0.00   => significant
#   Old-New  -2.76  -8.24  2.72  0.45   => not significant
#    Old-No   7.24   1.76 12.72  0.01   => significant
```


### effect size (practical significance) for the group differences

```{r}

# we need to subset!
# No vs New
cohen.d(PostTest ~ Condition,
        data = filter(surveydata, Condition != "Old"),
        paired = FALSE)

# No vs Old
cohen.d(PostTest ~ Condition,
        data = filter(surveydata, Condition != "New"),
        paired = FALSE)

# it makes no sense to calculate Cohen's d for Old-New, because it was not significant!

# NOTE: we are interest in the absolut value of Cohen's d (the the leading signs do not matter, they are arbitrary)
```


### visualize

```{r}
surveydata %>% 
  group_by(Condition) %>% 
  summarize(avg_PostTest = mean(PostTest)) %>% 
  ungroup() %>% 
  mutate(Condition = fct_reorder(Condition, avg_PostTest)) %>% 
  ggplot(aes(Condition, avg_PostTest,
             fill = Condition)) + 
  geom_col(show.legend = FALSE) + 
  labs(title = "Different Training Conditions",
       subtitle = "ANOVA",
       x = "Training Condition",
       y = "Average Post Test Score")
```

