---
title: "Centering & Standardizing Variables in R"
author: "Luca"
date: '2022-06-03'
output: html_document
---

Check out the full tutorial by David Caughlin [here](https://www.youtube.com/watch?v=yoPGwvUzjgQ&list=PLKkRkURCtPjCJOZHskCoyJCPb8wMDs2CW&index=23)


### load packages

```{r}
library(tidyverse)
library(psych)
library(lessR)

theme_set(theme_light())
```

### load data

```{r}
surveydata <- read_csv("https://raw.githubusercontent.com/davidcaughlin/R-Tutorial-Data-Files/master/DiffPred.csv")

surveydata
skimr::skim(surveydata)
```
 
 
 
### Centering: Grand-Mean Centering

better to interpret in an linear regression model for example or in the context of moderated linear regresssion (interaction terms) grand-mean centering is used before we create the interaction term in order to minimize co-linearity between the predictors and their product term


```{r}
# create the grand-mean centered interview value
surveydata <- surveydata %>% 
  mutate(centered_interview = interview - mean(interview, na.rm = TRUE))

mean(surveydata$interview,
     na.rm = TRUE)

# centered variable
mean(surveydata$centered_interview,
     na.rm = TRUE)
```



### standardizing

=> z-scores for each observation => mean = 0 and sd = 1!

```{r}
surveydata <- surveydata %>% 
  mutate(stand_interview = scale(.$interview),
         stand_interview = as.numeric(stand_interview))


mean(surveydata$interview)
sd(surveydata$interview)

# standardized variable
mean(surveydata$stand_interview)
sd(surveydata$stand_interview)
```


