---
title: "Chi-Square Test of Independence"
author: "Luca"
date: '2022-06-03'
output: html_document
---

Check out the full tutorial by David Caughlin [here](https://www.youtube.com/watch?v=yoPGwvUzjgQ&list=PLKkRkURCtPjCJOZHskCoyJCPb8wMDs2CW&index=23)


### load packages

```{r}
library(tidyverse)
library(psych)
library(lessR)

theme_set(theme_light())
```

### load data

```{r}
surveydata <- read_csv("https://raw.githubusercontent.com/davidcaughlin/R-Tutorial-Data-Files/master/ChiSquareTurnover.csv")

surveydata
skimr::skim(surveydata)
```
 
 
## Introduction

because of the data structure (2 variables of interest) we can run an 2x2 chi-squared independence test to check if the 2 variables are independt form each other.


### statistical assumptions
# 1. cases must be randomly sampled form the population
# 2. nonoccurences are included as well (here Stayed and Quit)



### create a contincency table (cross-tabulation) & run the chi-sqaured test (2x2)

```{r}
table <- xtabs(~ Onboarding + Turnover,
               data = surveydata)

table

# proportions instead of absolut values
# 2 means we look at the column proportions
prop.table(table, 2)

chisq.test(table)

# X-squared = 23.179, df = 1, p-value = 0.000001476
# => very low p-value => statistically significant association betweeen the two variables


# phi-coefficient
# => can be interpreted the same way as a Pearson's correlation (effect size metric)
phi(table)

# phi-coefficient = 0.59 (large effect)
```


### examine the ovsaerved and expected values

```{r}
chisq <- chisq.test(table)

chisq$observed
# what would be the expected values, if the two variables would truely independent of each other?
chisq$expected
# Pearson residuals (the larger the larger the influence of the chi-squared value)
chisq$residuals
```


