---
title: "<br>Structural Equation Model (SEM)<br>"
output: 
  html_document:
    toc: yes
    code_folding: show
editor_options: 
  chunk_output_type: console
---

<br>
<br>

# Structural Equation Models in R with lavaan
<br>
Source: https://clanfear.github.io/lavaan_tutorial/lavaan_tutorial_presentation.html#/
<br>
by Charles Lanfear (2018)

<br>
<br>

## Setup
```{r message=FALSE, warning=FALSE}
# Packages
library(tidyverse)
library(lavaan)
library(semPlot)
library(lavaanPlot)

# Data
data(HolzingerSwineford1939)
#?HolzingerSwineford1939
# Rename
df <- as_tibble(HolzingerSwineford1939)
```

<br>
<br>


# Linear Regression
```{r message=FALSE, warning=FALSE}
# with the lm()-function
lm_out_1 <- lm(x4 ~ ageyr,
               data = df)
summary(lm_out_1)

# with the sem()-function
sem_out_1 <- sem("x4 ~ ageyr",
                 data = df)
summary(sem_out_1)
```

<br>
<br>

# Multiple Regression
```{r message=FALSE, warning=FALSE}
# with the lm()-function
lm_out_2 <- lm(x4 ~ ageyr + sex + grade,
               data = df)
summary(lm_out_2)

# with the sem()-function
sem_out_2 <- sem("x4 ~ ageyr + sex + grade",
                 data = df)
summary(sem_out_2)
```

<br>
<br>

# Confirmatory Factor Analysis (CFA)
Here we have defined three latent (unobserved) variables, visual, textual, and speed, each with three observed indicators:
```{r message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
cfa_mod_1 <- "
visual =~ x1 + x2 + x3
textual =~ x4 + x5 + x6
speed =~ x7 + x8 + x9
"

cfa_out_1 <- cfa(cfa_mod_1,
                 data = df)
summary(cfa_out_1)

# Extract output as dataframe
parameterEstimates(cfa_out_1)

# Plot output
semPaths(cfa_out_1,
         whatLabels = "est")
### Does not work...
# lavaanPlot(model = cfa_out_1,
#            covs = TRUE, 
#            stand = FALSE))
```

<br>
**NOTE:**
<br>
When using the cfa() or sem() functions lavaan...
<br>
...Automatically sets the first indicator coefficient to 1
<br>
...Estimates covariances between latent variables
<br>
...Estimates variances for all latent and observed variables
<br>
...Defaults to a Maximum-Likelihood estimator.

<br>
<br>

# Structural Equation Model (SEM)

## 1. Walking dog
```{r message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
wd_mod_1 <- "
X =~ x1 + x2
Y =~ x4 + x5
Y ~ X
"
wd_out_1 <- sem(wd_mod_1,
                data = df)
summary(wd_out_1)
semPaths(wd_out_1,
         whatLabels = "est")
```

<br>

## 2. Fix particular coefficients 
We can fix particular coefficients to a certain value using pre-multiplication:
```{r message=FALSE, warning=FALSE}
### 2.
# we can fix particular coefficients to a certain value using 
# pre-multiplication
wd_mod_2 <- "
X =~ 1*x1 + 1*x2
Y =~ 1*x4 + 1*x5
Y ~ X
"
wd_out_2 <- sem(wd_mod_2,
                data =df)
summary(wd_out_2)
```

<br>

## 3. Force coefficients
We can force coefficients to have the same value:
```{r message=FALSE, warning=FALSE}
wd_mod_3 <- "
X =~ x1 + b1*x2
Y =~ x4 + b1*x5
Y ~ X
"
wd_out_3 <- sem(wd_mod_3,
                data = df)
summary(wd_out_3)
```

<br>

## 4.Constrain covariances
We can also constrain covariances to 0. Here we have made the assumption that the visual latent construct is orthogonal to textual and speed (aka has no influence):
```{r message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
wd_mod_4 <- "
visual =~ x1 + x2 + x3
textual =~ x4 + x5 + x6
speed =~ x7 + x8 + x9

visual ~~ 0*textual + 0*speed
"
wd_out_4 <- sem(wd_mod_4,
                data =df)
summary(wd_out_4)
semPaths(wd_out_4,
         whatLabels = "est")
```

<br>

## 5. Complex models
We can build complex models (multiple latent variables, regressions and covariances in one model):
```{r message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
complex_mod_1 <- "
visual =~ x1 + x2 + x3
textual =~ x4 + x5 + x6
speed =~ x7 + x8 + x9

visual ~ textual + speed
textual ~ speed
"
complex_out_1 <- sem(complex_mod_1,
                     data = df)
summary(complex_out_1)
parameterestimates(complex_out_1)
semPaths(complex_out_1,
         whatLabels = "est")
```

<br>

## 6. Alternate estimaors
lavaan defaults to using a full-information Maximum-Likelihood (ML) Estimator. This is the most efficient estimator available and can handle missing data (missing=ml). We may want to specify alternate estimators if our data do not meet the assumptions for ML. By adding the estimator= argument to our cfa(), sem(), or lavaan() call, we can choose another estimator.
<br>
**Some useful choices:**
<br>
**estimator="MLM":** ML with robust errors and Satorra-Bentler test statistic.
<br>
**estimator="WLSMV":** Weighted least squares with mean and variance adjusted test statistic (needed for categorical endogenous variables).

<br>

## 7. Categorical variables
If you provide lavaan categorical data as an endogenous variable, it will automatically use a proper estimator (e.g. Diagonally Weighted Least Squares). To specify an ordinal variable as categorical, you will want to make it an ordered factor:
```{r message=FALSE, warning=FALSE}
iris %>% 
  as_tibble() %>% 
  mutate(Species = ordered(Species)) %>% 
  glimpse()
```

Unordered categorical variables (factors) will need to be split into dummy variables prior to estimation:
```{r message=FALSE, warning=FALSE}
iris %>% 
  fastDummies::dummy_cols(select_columns = "Species",
                          remove_first_dummy = TRUE,
                          remove_selected_columns = TRUE) %>% 
  glimpse()
```


