---
title: "Cyclotour 2023"
output: 
  html_document:
    toc: yes
    code_folding: show
---
<br>
<br>

#### Load Packages

```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(ggrepel)
library(glue)
library(countrycode)
library(sjPlot)

# set theme and figure size
theme_set(theme_minimal())
knitr::opts_chunk$set(fig.width=14, 
                      fig.height=10)
```

<br>

#### Load data

```{r warning=FALSE, message=FALSE}
df <- read_rds("Cyclotour_2023/cyclotour.rds") %>% 
  mutate(age = 2023 - year_of_birth,
         club = ifelse(team == "",
                       "No Club",
                       "Club")) %>% 
  filter(age <= 100) %>% 
  left_join(codelist %>% 
              select(country.name.en, continent, region23, ioc),
            by = c("country" = "ioc")) %>% 
  relocate(start_number, name, year_of_birth, 
           age, continent, country, "country_name" = country.name.en, 
           "region" = region23, city, team, club, bike_type, time)
```

<br>

## Descriptive statistics

### How old are the participants?

```{r warning=FALSE, message=FALSE}
df %>% 
  ggplot(aes(age)) + 
  geom_histogram(color = "white",
                 fill = "seagreen",
                 alpha = 0.75) + 
  
  # WE
  geom_vline(xintercept = 28,
             lty = 2, 
             color = "red",
             size = 2) + 
  annotate(geom = "label",
           x = 28, 
           y = 175, 
           label = "Andy\nArvid\nLuca") +
  
  # Average
  geom_vline(xintercept = mean(df$age, na.rm = TRUE),
             linewidth = 1.25,
             alpha = 0.5) +
  annotate(geom = "label",
           x = mean(df$age, na.rm = TRUE),
           y = 175,
           label = "Average") +
  
  annotate(geom = "label",
           x = 75, 
           y = 150, 
           label = glue("min: {min(df$age, na.rm = TRUE)}
                        Q25: {quantile(df$age, 0.25, na.rm = TRUE)}
                        Q50: {quantile(df$age, 0.50, na.rm = TRUE)}
                        avg: {round(mean(df$age, na.rm = TRUE), 1)}
                        Q75: {quantile(df$age, 0.75, na.rm = TRUE)}
                        max: {max(df$age, na.rm = TRUE)}")) +
  labs(title = "\nHow old are the participants?",
       subtitle = glue("Total: {nrow(df)} Participants\n"),
       x = "\nAge (years)",
       y = "Count\n",
       caption = "Outliers (age >= 100) have been removed.") + 
  scale_x_continuous(breaks = seq(15, 80, 5))
```

<br>

### Where did the participants come from?

```{r warning=FALSE, message=FALSE}
world_map <- map_data('world') %>% 
  filter(region != "Antarctica")


# create subset for country labels
country_labels <- world_map %>% 
  group_by(region) %>%
  mutate(long = mean(long), lat = mean(lat)) %>% 
  select(c("region", "long", "lat")) %>% 
  unique()

# US fix
country_labels <- country_labels %>% 
  mutate(long = ifelse(region == "USA",
                       -105,
                       long),
         lat = ifelse(region == "USA",
                      40,
                      lat)) 

# add longitude and latitude to data
df <- df %>% 
  left_join(country_labels,
            by = c("country_name" = "region"))

# map
world_map %>% 
  ggplot() + 
  geom_polygon(aes(long, lat,
                   group = group),
               color = "grey90",
               fill = "grey80") + 
  geom_point(aes(long, lat,
                 color = n,
                 size = n),
             data = df %>% 
               add_count(country_name)) + 
  geom_text_repel(aes(x = long, y = lat, 
                      label = region),
                  color = "black", 
                  size = 3,
                  max.overlaps = Inf,
                  data = country_labels %>% 
                    filter(region %in% df$country_name)) +
  scale_color_gradient2(low = "blue", 
                        mid = "orange",
                        high = "red",
                        midpoint = 750) +
  scale_size(range = c(1, 8)) +
  coord_map(projection = "mercator",
            xlim = c(-180, 180)) +
  labs(title = "\nWhere did the participants come from?",
       subtitle = glue("Total: {nrow(df)} Participants\n"),
       x = "\nLongitude",
       y = "Latitude\n",
       color = "Number of\nparticipants:",
       caption = "Outliers (age >= 100) have been removed.") + 
  guides(size = "none")
```

<br>

### How did WE do overall?

```{r warning=FALSE, message=FALSE}
df %>% 
  ggplot(aes(time)) +
  geom_histogram(color = "white",
                 fill = "seagreen",
                 alpha = 0.75) +
  
  annotate(geom = "label",
           x = 8.75,
           y = 175,
           label = glue("min: {round(min(df$time, na.rm = TRUE), 2)}
                        Q10: {round(quantile(df$time, 0.10, na.rm = TRUE), 2)}
                        Q25: {round(quantile(df$time, 0.25, na.rm = TRUE), 2)}
                        Q50: {round(quantile(df$time, 0.50, na.rm = TRUE), 2)}
                        avg: {round(mean(df$time, na.rm = TRUE), 2)}
                        Q75: {round(quantile(df$time, 0.75, na.rm = TRUE), 2)}
                        Our Team: {round(7.038889, 2)}
                        Q90: {round(quantile(df$time, 0.90, na.rm = TRUE), 2)}
                        max: {round(max(df$time, na.rm = TRUE), 2)}")) + 
  # WE
  geom_vline(xintercept = 7.038889,
             lty = 5,
             linewidth = 2,
             color = "red") +
  annotate(geom = "label",
           x = 7.038889, 
           y = 185, 
           label = "Andy\nArvid\nLuca") +
  
  # Top 10%
  geom_vline(xintercept = quantile(df$time, 0.10, 
                                   na.rm = TRUE),
             linewidth = 1.25,
             alpha = 0.5) +
  annotate(geom = "label",
           x = quantile(df$time, 0.10, 
                        na.rm = TRUE), 
           y = 150, 
           label = "Top 10%") +
  
  # Top 25%
  geom_vline(xintercept = quantile(df$time, 0.25, 
                                   na.rm = TRUE),
             linewidth = 1.25,
             alpha = 0.5) +
  annotate(geom = "label",
           x = quantile(df$time, 0.25, 
                        na.rm = TRUE), 
           y = 125, 
           label = "Top 25%") +
  
  # Top 50%
  geom_vline(xintercept = quantile(df$time, 0.5, 
                                   na.rm = TRUE),
             linewidth = 1.25,
             alpha = 0.5) +
  annotate(geom = "label",
           x = quantile(df$time, 0.5, 
                        na.rm = TRUE), 
           y = 150, 
           label = "Top 50%") +
  
  # Top 75%
  geom_vline(xintercept = quantile(df$time, 0.75, 
                                   na.rm = TRUE),
             linewidth = 1.25,
             alpha = 0.5) +
  annotate(geom = "label",
           x = quantile(df$time, 0.75, 
                        na.rm = TRUE), 
           y = 125, 
           label = "Top 75%") +
  
  # Top 90%
  geom_vline(xintercept = quantile(df$time, 0.90, 
                                   na.rm = TRUE),
             linewidth = 1.25,
             alpha = 0.5) +
  annotate(geom = "label",
           x = quantile(df$time, 0.90, 
                        na.rm = TRUE), 
           y = 150, 
           label = "Top 90%") +
  
  scale_x_continuous(breaks = seq(4, 10, 0.25)) +
  labs(title = "\nHow did WE do overall?",
       subtitle = glue("Total: {nrow(df)} Participants\n"),
       x = "\nTime (hours)",
       y = "Count\n",
       caption = "Outliers (age >= 100) have been removed.")
```

<br>

### Age-Effect?

```{r warning=FALSE, message=FALSE}
df %>% 
  ggplot(aes(age, time)) + 
  geom_jitter() +
  geom_smooth(method = "lm") +
  labs(title = "\nAge-Effect?",
       subtitle = glue("Total: {nrow(df)} Participants\n"),
       x = "\nAge (years)",
       y = "Time (hours)\n",
       caption = "Outliers (age >= 100) have been removed.") + 
  scale_x_continuous(breaks = seq(15, 80, 5)) + 
  scale_y_continuous(breaks = seq(4, 9, 1))
```

=> Maybe a tiny linear age-effect (the older a person the more time is needed).

<br>

### Bike-Type-Effect?

```{r warning=FALSE, message=FALSE}
df %>% 
  add_count(bike_type) %>% 
  mutate(bike_type = fct_reorder(bike_type, -time),
         bike_type = str_to_title(bike_type),
         bike_type = glue("{bike_type} (n = {n})")) %>% 
  ggplot(aes(bike_type, time)) + 
  geom_boxplot(aes(fill = bike_type),
               show.legend = FALSE) +
  scale_y_continuous(breaks = seq(4, 10, 0.5)) +
  labs(title = "\nBike-Type-Effect?",
       subtitle = glue("Total: {nrow(df)} Participants | Number of Observations in (...)\n"),
       x = "",
       y = "Time (hours)\n",
       caption = "Outliers (age >= 100) have been removed.")
```

=> Electric bikes are the slowest. Probably because mainly older people use them.
<br>
NOTE: Not enough observations for tandem and electric bikes...

<br>

### Club-Effect?

```{r warning=FALSE, message=FALSE}
df %>% 
  add_count(club) %>% 
  mutate(club = glue("{club} (n = {n})")) %>% 
  ggplot(aes(club, time)) + 
  geom_boxplot(aes(fill = club),
               show.legend = FALSE) +
  scale_y_continuous(breaks = seq(4, 10, 0.5)) +
  labs(title = "\nClub-Effect?",
       subtitle = glue("Total: {nrow(df)} Participants | Number of Observations in (...)\n"),
       x = "",
       y = "Time (hours)\n",
       caption = "Outliers (age >= 100) have been removed.")
```

=> On average people who are in a bicycle-club are faster. 
<br>
<br>
But which factors are <b>statistically significant?</b> 

<br>

## Linear Regression Models

```{r warning=FALSE, message=FALSE}
df_model <- df %>% 
  mutate(time = time * 60,
         club = as.factor(club),
         bike_type = as.factor(bike_type),
         continent = as.factor(continent))

mod_1 <- lm(time ~ age, 
            data = df_model)

mod_2 <- lm(time ~ age + club,
            data = df_model)

mod_3 <- lm(time ~ age + club + bike_type,
            data = df_model)

mod_4 <- lm(time ~ age + club + bike_type + continent,
            data = df_model)

tab_model(mod_1, mod_2, mod_3, mod_4,
          dv.labels = c("Model 1", "Model 2",
                        "Model 3", "Model 4"))
```
<br>
<b>All factors</b> are statistically significant. <b>Model 3</b> has the most explanatory power ($R^{2} adjusted$ = 0.026 ~ 2.6% of the variance can be explained by the model, which is nothing...)

<b>AGE:</b> If the age of a person increases by one year, the time increases by 0.44 minutes (~26.5 seconds) on average.
<br>
<b>CLUB:</b> If you are in a cycling club, you are 17.5 minutes faster on average.
<br>
<b>BIKE TYPE:</b> With an electric bike you are on average 77.4 (85.5 minutes) slower than with a normal bike (tandem bike).
<br>
<br>
<b>NOTE:</b> Not enough cases in some groups (bike type, continent). The predictions are not to be taken too seriously...
